Acceleration of Derivative Calculations with Application to Radial Basis Function -- Finite-Differences on the Intel MIC Architecture

### 2013-10-19 ieee_paper.tex
another mbox to prevent hyphenation

Loading the entries from the {\tt x} vector is a little bit more
involved. First, the column pointers of four different nonzeros are
loaded into a vector register and the values are distributed one per
lane of the vector register using an {\tt unpack}
operation\footnote{Notice that there are two calls to {\tt unpack} for the
LSB and MSB. Both are mandatory even if one is known to be
nilpotent according the documentation of the hardware.} (The column
pointers are named $j,k,l,m$ in Figure~\ref{code:mat_mul}.) They
are replicated so that each value fills a single vector lane 
using a swizzle operation. Then each value is multiplied by four
(because $n_v=4$) 
and the offset vector $(3,2,1,0)$ is added to each lane to obtain the
correct index of the elements of the {\tt x} vector within each lane
using a Fused Multiply-Add operation. A {\tt gather} operation is then
performed to bring the 16 entries of the {\tt x} vector into the SIMD
register. At this point we have a register where each lane is the
four vector entries for a nonzero element. Using a broadcast
permutation, one can replicate an entry of the four vectors across
each lane. (The broadcast of $x_m^1,x_m^2,x_m^3,x_m^4$ is shown in
Figure~\ref{code:mat_mul}.)

@@

In this section, we describe a series of numerical experiments to
confirm the performance predictions of Section~\ref{sec:model}. To
better understand our results, we first perform some bandwidth
measurements, followed by the actual SpMV computations. All the codes
are compiled with the Intel C Compiler version 13 and
-O3 optimization. The codes use the IMCI instructions exposed by
the compiler through intrinsics.

@@

Figure~\ref{fig:bandwidth} shows a benchmark that computes the
bandwidth achieved when reading and/or writing large arrays on both
the MIC card and the CPU architecture. The read benchmark performs a
simple sum of the array. The write benchmark sets all the elemnts of
the array to zero, while the read/write benchmarks copies the array into
another one. We also investigate the difference between a straight C++
implementation (compiled with -O3 with proper tagging to inform the
compiler of memory alignments) and a simple implementation with
explicit use of the vector registers.

>>>>

Loading the entries from the {\tt x} vector is  more
involved. First, the column pointers of four different nonzeros are
loaded into a vector register and the values are distributed one per
lane of the vector register using an {\tt unpack}
operation\footnote{Notice that there are two calls to {\tt unpack} for the
LSB and MSB. Both are mandatory even if one is known to be
nilpotent according the documentation of the hardware.} (The column
pointers are named $j,k,l,m$ in Figure~\ref{code:mat_mul}.) They
are replicated so that each value fills a single vector lane 
using a swizzle operation. Then each value is multiplied by four
(because $n_v=4$) 
and the offset vector $(3,2,1,0)$ is added to each lane to obtain the
correct index of the elements of the {\tt x} vector within each lane
using a Fused Multiply-Add operation. A {\tt gather} operation is then
performed to bring the 16 entries of the {\tt x} vector into the SIMD
register. At this point we have a register where each lane is the
four vector entries for a nonzero element. Using a broadcast
permutation, one can replicate an entry of the four vectors across
each lane. (The broadcast of $x_m^1,x_m^2,x_m^3,x_m^4$ is shown in
Figure~\ref{code:mat_mul}.)

@@

In this section, we describe a series of numerical experiments to
confirm the performance predictions of Section~\ref{sec:model}. To
better understand our results, we first perform some bandwidth
measurements, followed by the actual SpMV computations. All the codes
are compiled with the Intel C Compiler version 13 and
\mbox{-O3} optimization. The codes use the IMCI instructions exposed by
the compiler through intrinsics.

@@

Figure~\ref{fig:bandwidth} shows a benchmark that computes the
bandwidth achieved when reading and/or writing large arrays on both
the MIC card and the CPU architecture. The read benchmark performs a
simple sum of the array. The write benchmark sets all the elemnts of
the array to zero, while the read/write benchmarks copies the array into
another one. We also investigate the difference between a straight C++
implementation (compiled with \mbox{-O3} with proper tagging to inform the
compiler of memory alignments) and a simple implementation with
explicit use of the vector registers.
